---
title: "Knockoff Filter"
author: "Michele Filopanti, Elena Maggiore, Marco Rosato"
date: "2025-11-22"
output:
  word_document: default
  pdf_document: default
---

```{r setup, echo=F}
library(glmnet)
library(caret)
library(mice)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(e1071) 
library(gridExtra)
library(pROC)
library(car)
library(corrplot)
library(knockoff)
library(ranger)

knitr::opts_chunk$set(echo = TRUE)

na.fill <- function(col) {
  is.num <- is.numeric(col)
  col[is.na(col)] <- ifelse(is.num, 0, "None")
  if(is.num) col else as.factor(col)
}
```

## Importing the data

Data is found at https://www.kaggle.com/datasets/shashanknecrothapa/ames-housing-dataset. We will mostly consider the numeric columns, as categorical columns are not suitable for the construction of knockoff variables. However, categorical columns will still prove useful in the preprocessing phase.

```{r import}
df0 <- read.csv("AmesHousing.csv") |> dplyr::select(-Order, -PID) |> Filter(f= function(x)length(unique(x)) > 0)
colnames(df0) <- gsub("\\.", "",  colnames(df0))
df <- df0  |> select_if(is.numeric) |> select(-MSSubClass, -TotalBsmtSF)
summary(df)[, 1:10]
```

It is immediately visible that some columns present missing values. Here are the columns containing missing values, with the respective count:

```{r na.count}
df |> sapply(function(x)sum(is.na(x))) |> Filter(f=identity) -> na.counts
print(na.counts)
```
The `LotFrontage` variable is missing in a sizeable portion of the records, and would therefore be unwise to remove. Same goes for `GarageYrBlt`. The other variables contain a minimal amount of missing values. most of them in the same row, meaning we can safely discard the incomplete records.

```{r}
na.counts |> Filter(f = function(x)x < 100) |> names() -> cols.to.check
rows.to.keep <- complete.cases(df[, cols.to.check])
df <- df[rows.to.keep, ]
df0 <- df0[rows.to.keep, ]
```

If we inspect the values of the other garage-related variables when `GarageYrBlt` is missing, we can see that these missing values are not at random:

```{r GarageYrBltNAs}
df[is.na(df$GarageYrBlt), c("GarageYrBlt", "GarageCars", "GarageArea")]
```

meaning that a missing `GarageYrBlt` value is not due to registration errors, but rather signifying a house with a missing garage. We will therefore fill these NA values with zeros.

```{r GarageYrBltNAs.2}
df <- transform(df, GarageYrBlt = na.fill(GarageYrBlt))
```

Regarding `LotFrontage`, can notice one thing: there are some other lot-related variables in the dataset, those being `LotShape` (categorical), `LotConfig` (categorical) and `LotArea` (numeric). Together with `Neighbourhood` and `BldgType` (both categorical), they seem to explain decently well the missingness of `LotFrontage`.

We can see that houses whose `LotShape` is regular are much less likely to have a missing `LotFrontage`:

```{r}
cols <- c("LotShape", "LotFrontage", "LotConfig", "LotArea", "Neighborhood", "BldgType")
df.lot <- df0[cols]

with(df.lot, table(is.na(LotFrontage), LotShape)) |> apply(2, function(x) x / sum(x))
```

therefore, before modeling the missingness, we will group together all irregular lot shapes in a single level. We will also merge all `Neighborhood` levels which have 0 missing response values with the level with the least amount of missing values.

```{r}
sapply(df.lot$LotShape, function(x)ifelse(x != "Reg", "Irreg", x)) -> df.lot$LotShape

with(df.lot, table(is.na(LotFrontage), Neighborhood)) |> apply(2, function(x) x / sum(x)) -> ngbr.table
ngbr.table
ngbr.table[1, ] |> sort(decreasing = T) |> names() -> to.fuse
to.fuse <- to.fuse[1:3]
df.lot$Neighborhood <- sapply(df.lot$Neighborhood, function(x)ifelse(x %in% to.fuse, "Other", x)) |> as.factor() |> relevel(ref = "Other")
```

From the logistic model below...

```{r}
df.lot$LotFrontage |> is.na() |> as.factor() -> df.lot$LotFrontage

lot.model <- glm(LotFrontage ~ ., data = df.lot, family = binomial())
lot.roc <- roc(df.lot$LotFrontage, fitted(lot.model))
plot.roc(lot.roc)
lot.roc$auc
```
... we can see that the selected variables appear to explain decently well the missingness of `LotFrontage`.
By checking the coefficients effects of each variable, we can see that lots with a regular shape indeed are much more likely to not have a missing lot frontage value. Also, inside lots and cul-de-sac lots are more prone to missingness.

```{r}
summary(lot.model)
```

Missing data imputation will proceed in the following way:
* for columns with **irregular** lot shape, the lot frontage will be imputed with 0
* for columns with **regular** lot shape, imputation using a random forest will be performed

```{r}
regs <- which(df0$LotShape != "Reg" & is.na(df0$LotFrontage))
df[regs, "LotFrontage"] <- df0[regs, "LotFrontage"] <- 0

for(col in colnames(df0)[c(1,2,4:80)])
  df0[[col]] <- na.fill(df0[[col]])
```

Before proceeding with model-based missing data imputation, we ought to divide the dataset in train and test, as to not contaminate the test data and preserve its independence with the training data.

```{r}
rm(list = ls(all = F) |> setdiff(c("df", "df0")))
set.seed(42)

train.rows <- sample(1:nrow(df), replace = F, size = .7 * nrow(df))
df.train <- df[train.rows, ]
df.test <- df[-train.rows, ]

dummyVars(~ ., data = df0) |> predict(newdata = df0) |> data.frame() |> select(-SalePrice) -> data
data.train <- data[train.rows, ]
data.test <- data[-train.rows, ]

data.train.fit <- data.train[!is.na(data.train$LotFrontage), ]
data.train.imp <- data.train[is.na(data.train$LotFrontage), ]
data.test.imp <- data.test[is.na(data.test$LotFrontage), ]

ctrl <- trainControl(method = "cv", number = 5, search = "grid")
rf <- train(LotFrontage ~ ., data = data.train.fit, method = "rf", 
            tuneGrid = expand.grid(.mtry = 5:7), ntree = 100, trControl = ctrl) 

train.imp <- predict(rf, newdata = data.train.imp) 
test.imp <- predict(rf, newdata = data.test.imp) 
```

```{r}
df.train[is.na(df.train$LotFrontage), "LotFrontage"] <- train.imp
df.test[is.na(df.test$LotFrontage), "LotFrontage"] <- test.imp

rm(list = ls(all = F) |> setdiff(c("df.train", "df.test")))
```

# Target variable

```{r plot.y}
p1 <- ggplot(df.train, aes(x=SalePrice)) + 
  geom_histogram(fill="red", bins=100) + ggtitle("SalePrice Distribution")
p1
```

We can observe that the target variable is right skewed: we will check the most appropriate Box-Cox transformation to make it more normally distributed.

```{r boxcox}
car::boxCox(lm(SalePrice ~ ., data = df.train))
```
The optimal value for `r expression(lambda)` appears to be 0: we will therefore apply a logarithmic transformation. What follows is a before-after comparison:

```{r plot2}
p2 <- ggplot(df.train, aes(x=log(SalePrice))) + 
  geom_histogram(fill="blue", bins=100) + ggtitle("Log(SalePrice) Distribution")

grid.arrange(p1, p2, ncol=2)

df.train$SalePrice <- log(df.train$SalePrice)
df.test$SalePrice <- log(df.test$SalePrice)
```


# Non zero variance variables

```{r}

near_zero_var <- nearZeroVar(df.train, freqCut = 19, uniqueCut = 10,allowParallel = TRUE, saveMetrics = TRUE) 
near_zero_var 

df.train <- df.train[, !near_zero_var$nzv] 
df.test <- df.test[, !near_zero_var$nzv]
dim(df.train)

```


# Correlation Analysis

In the following section, we will check the raw correlation and partial between the predictors.

```{r corr, fig.height = 10}
X_raw <- df.train[, names(df.train) != "SalePrice"]
y_raw <- df.train$SalePrice

corrplot::corrplot(cor(X_raw))
```

We can see clusters of variables that are more correlated, particularly variables **garage** ("GarageArea", "GarageCars", "GarageYrBlt"). Also, one can notice how variables "OverallQlt" and "GrLivArea" appear to be slightly correlated with almost all other predictors. 


# Matrix rank

Before continuing, we want to be sure that our matrix X is full rank, and remove linear dependent columns otherwise.
```{r}
rankMatrix(X_raw)
dim(X_raw)
X_clean=X_raw
```
Our matrix is full rank.


# High correlation check

Now, we check for highly correlated columns, removing all variables that are shown to be more correlated to another predictor with a correlation stronger or equal than 0.8 (absolute value).
We also obtain the Variance Inflation Factor for each variable, to better understand how correlation amongst predictors adds to the variance.

```{r}
vif_values <-vif(lm(y_raw~.,data=X_clean))
print(vif_values[vif_values>10])

cor_matrix_clean <- cor(X_clean)
high.corrs <- findCorrelation(cor_matrix_clean, cutoff=0.8)
high.corrs <- colnames(X_clean[high.corrs])
print(high.corrs)

if(length(high.corrs) > 0){
  X_clean <- X_clean |> select(-all_of(high.corrs))
  df.train <- df.train |> select(-all_of(high.corrs))
  df.test <- df.test |> select(-all_of(high.corrs))
}
```

Our matrix has 2032 rows and 25 columns.


# Outlier detection

We will search for outliers using Cook Distance, and remove rows that appear potentially problematic.

```{r}
lm_cook <- lm(y_raw ~ ., data = X_clean)
cookdis <- cooks.distance(lm_cook)

plot(cookdis, type="h", main="Cook's Distance", ylab="Cook's Distance")
abline(h = 4/(nrow(X_clean)-ncol(X_clean)), col="red")

outlier_indices <- which(cookdis > (4/(nrow(X_clean)-ncol(X_clean))))
outlier_indices

X_clean <- X_clean[-outlier_indices, ]
dim(X_clean)
y_clean <- y_raw[-outlier_indices]
```

Now, matrix X has 1937 rows and 25 columns.


# Centering and Scaling

To use the Knockoff Technique, we need our data to be centered and scaled.

```{r}
normalize <- function(data) as.matrix(data) |> scale() |> apply(2, function(t) t / sqrt(length(t) - 1))

X <- normalize(X_clean)
y <- normalize(y_clean)

summary(X)
```


# Hypothesis check

Finally, we must ensure the following hypothesis to be satisfied.

* The sum of squares should be 1:

```{r}
unique(apply(X, 2, function(col) sum(col^2)))
```

* Means should be 0:

```{r}
unique(apply(X, 2, function(col) round(sum(col), 10)))
```

* The Fixed-X Method requires that:

```{r}
nrow(X) >= 2*ncol(X)
```

* Finally, matrix X must be full rank:

```{r}
rankMatrix(X)
dim(X)
```

This guarantees that we can use the Fixed-X approach of the Knockoff Filter.


# Fixed-X Knockoff

Now, we're going to perform the Fixed-X Knockoff technique, and estimate the False Discovery Rate after selecting the most significant predictors through a Lasso selection.

```{r}
rm(list = ls(all = F) |> setdiff(c("df.train", "df.test" , "X", "y")))

summary(X)
summary(y)

n <- nrow(X)
p <- ncol(X)
varType <- rep("N",p)
```


# Knockoff construction

First, we obtain matrix Sigma and its inverse.

```{r}
Sigma <- crossprod(X)
Sigma_inv <- solve(crossprod(X))
```

We construct U so that U'U = I_p and U'X = 0_p. In order to do so, we conjoin X by row with a zero-matrix of the same dimensions, obtaining X_0 of size (nx2p). 

By computing the SVD of X_0 and taking the last p columns of the obtained U matrix, we get a matrix that is orthonormal by construction, and whose column space is orthogonal to the column space of X (which coincides with the column space of the first p columns of U).

```{r}
X_svd <- X |> cbind(matrix(0,n,p)) |> svd()
U <- X_svd$u[,(p+1):(2*p)]
```

Now that we have Sigma, Sigma_inv and U, we can create knockoff variables.
First, we'll use the Equi-Correlated Knockoffs method. To obtain the maximum possible value of d, we use the formula:

```{r}
d <- min(c(2 * svd(X)$d^2, 1)) * (1 - 1e-7)
d

d <- diag(rep(d, p))

C <- chol(d %*% (2 * diag(rep(1, p)) - Sigma_inv %*% d))
Xtilde <- X %*% (diag(rep(1, p)) - Sigma_inv %*% d) + U %*% C
```

We obtain d = 0.1496.

Let's check the first three rows and columns to see if matrix G is correctly constructed:

```{r}
crossprod(X)[1:3,1:3]
crossprod(Xtilde)[1:3,1:3]
crossprod(X,Xtilde)[1:3,1:3]

round(crossprod(X)[1:3,1:3],5)==round(Sigma[1:3,1:3],5)
round(crossprod(Xtilde)[1:3,1:3],5)==round(Sigma[1:3,1:3],5)

round(crossprod(X,Xtilde)[1:3,1:3],5)==round(Sigma[1:3,1:3]-diag(rep(d,p))[1:3,1:3],5)

```


# Plots about knockoff variables

These plots show the correlation matrix of X and the correlation matrix of Xtilde, which should be identical. The third plot shows the correlation matrix between X and Xtilde (note that the diagonal represents 1-d).

```{r}
par(mfrow = c(1,3))
plot1 <- corrplot(crossprod(X), 
         method = "color", 
         order = "hclust",
         tl.pos = "n", 
         addgrid.col = NA,
         title = "Correlation Matrix of X", 
         mar = c(0,0,1,0))

plot2 <- corrplot(crossprod(Xtilde), 
         method = "color", 
         order = "hclust",
         tl.pos = "n", 
         addgrid.col = NA,
         title = "Correlation Matrix of Xtilde", 
         mar = c(0,0,1,0))


plot3 <- corrplot(crossprod(X,Xtilde), 
         method = "color", 
         order = "hclust",
         tl.pos = "n", 
         addgrid.col = NA,
         title = "Correlation Matrix of X & Xtilde", 
         mar = c(0,0,1,0))
```

This plot will show us correlation between each pair of variables X_i and Xtilde_i.

```{r}
cor_cols <- sapply(seq_len(ncol(X)), function(i) {
  x <- X[, i]
  xtilde <- Xtilde[, i]
  cor(x, xtilde, use = "pairwise.complete.obs")
})

par(mfrow = c(1,1))
plot(cor_cols, type = "h", lwd = 2,
     xlab = "Index i", ylab = "Correlation",
     main = "Correlation for each pair i of X and Xtilde",
     ylim = c(-1, 1))
points(cor_cols, pch = 16)
abline(h = 0, lty = 2)
```

This result depends on the fact that we're using the Equi-Knockoff method.


# Knockoff Statistic

We combine X and Xtilde, fit a Lasso using glmnet and find the best lambda through cross-validation.

```{r}
XX <- cbind(X,Xtilde)
fit <- glmnet(XX, y)
l <- (cv.glmnet(XX, y)$lambda.min)
```

This plot shows the estimates of the coefficients; blue variables are knockoff variables.

```{r}
cols <- c(rep(1, p), rep(4, p))   
plot(1:(2*p), coef(fit, s=l)[-1], 
     xlab="Index j", 
     ylab=expression(hat(beta)[j]), 
     pch=19,
     col=cols)
```

We want to find out which of the selected variables are knockoffs (K) and which are actual columns of X (N). The Lasso should pick as few knockoff variables as possible. We won't consider the intercept.

```{r}
varType <- c(rep("N",p),rep("K",p))
S_hat <- which(coef(fit, s=l)[-1] != 0)
print(S_hat)
table(varType[S_hat])
```

Now, we calculate the Knockoff Statistic for each pair of variables i.
First, we find the first lambda for which a variable enters the model (if it's null, we put 0), then calculate W statistic for each pair of variables.

We can calculate Z:
```{r}
first_nonzero <- function(x) match(T, abs(x) > 0)
indices <- apply(fit$beta, 1, first_nonzero)
Z = ifelse(is.na(indices), 0, fit$lambda[indices]* n)
Z
```

And then W:

```{r}
orig = 1:p
W = pmax(Z[orig], Z[orig+p]) * sign(Z[orig] - Z[orig+p])
W
```


# FDP Estimate

First, let's see an example where we choose tau=2:

```{r}
tau = 2

my_cols <- ifelse(W >= tau, "red", "black")
plot(Z[orig], Z[orig+p], col=my_cols, pch=19, asp=1, 
     xlab=expression(paste(lambda," when ",X[j]," enters ")),
     ylab=expression(paste(lambda," when ",tilde(X)[j]," enters ")),
     main=paste("Knockoff Selection (tau =", tau, ")"))
abline(a=0,b=1)
abline(h=tau,lty=3)
abline(v=tau,lty=3)
legend("topleft", legend=c("Selected", "Not Selected"), col=c("red", "black"), pch=19)
```

The model selects all variables on the right of the vertical line (tau=2) and below the diagonal (positive W).

```{r}
S_hat_tau = which(W >= tau)
S_hat_tau
table(varType[S_hat_tau])

(1 + sum(W <= -tau)) / length(S_hat_tau)
```

These are the 12 selected variables.
By putting tau=2, we obtain an FDP estimate of 0.25.

To control our estimate of the FDP, we can chose alpha as the maximum tolerable error; then, the algorithm automatically chooses the best value of tau such that the FDP estimate is lower than alpha.

For each possible tau, we find the expected FDP.

```{r}
taus = sort(c(0,abs(W)))

FDP_hat = sapply(taus, function(tau)
  (1 + sum(W <= -tau)) / sum(W >= tau))
FDP_hat
```

Now, we can choose a value of alpha and find the best tau for that value.

```{r}
alpha = 0.1
tau_hat <- taus[[which.min(FDP_hat[FDP_hat  <= alpha]) |> names()]]

print(paste("Selected tau:", tau_hat))
S_hat = which(W >= tau_hat)
print(paste("Estimated FDP:", round((1 + sum(W <= -tau_hat)) / length(S_hat), 3)))
```

The selected tau is 4.3281, which generates an estimated FDP of 0.091.

```{r}
equi.vars <- colnames(X)[which(W >= tau_hat)]
table(varType[S_hat])
print(equi.vars)
```

We selected the following 10 variables.

This plot shows our results.

```{r}
plot(taus, FDP_hat, type="l", lwd=2, 
     xlab=expression(tau), ylab="FDP Estimate",
     main="Selecting the Optimal Threshold (Knockoff)")
abline(h = alpha, col="red", lty=2)
abline(v = tau_hat, col="blue", lwd=2)
text(x=tau_hat, y=0.5, labels=paste("Selected Tau =", round(tau_hat, 3)), pos=4, col="blue")
```


# Knockoff SDP

We could also use Semidefinite-Programming Knockoffs by using function *create.fixed* from package **knockoffs**.

```{r}
Xtilde_sdp <- create.fixed(X,method="sdp")$Xk
```

We observe the correlation between each pair of variables from X and X_tilde, now using SDP-Knockoffs.

```{r}
cor_cols <- sapply(seq_len(ncol(X)), function(i) {
  x <- X[, i]
  xtilde <- Xtilde_sdp[, i]
  cor(x, xtilde, use = "pairwise.complete.obs")
})

par(mfrow = c(1,1))
plot(cor_cols, type = "h", lwd = 2,
     xlab = "Index i", ylab = "Correlation",
     main = "Correlation for each pair i of X and Xtilde",
     ylim = c(-1, 1))
points(cor_cols, pch = 16)
abline(h = 0, lty = 2)
```
Now, correlation between X and Xtilde is different for each variable.

We repeat the process of fitting a Lasso and estimating the Knockoff Statistic W, using the SDP-Knockoffs:

```{r}
XX_sdp <- cbind(X, Xtilde_sdp)

fit_sdp <- glmnet(XX_sdp, y)

indices_sdp <- apply(fit_sdp$beta, 1, first_nonzero)

Z_sdp = ifelse(is.na(indices_sdp), 0, fit_sdp$lambda[indices_sdp] * n)

W_sdp = pmax(Z_sdp[orig], Z_sdp[orig+p]) * sign(Z_sdp[orig] - Z_sdp[orig+p])
```

And, for alpha=0.1, find the value of tau and the best estimate of FDP.

```{r}
taus_sdp = sort(c(0, abs(W_sdp)))

FDP_hat_sdp = sapply(taus_sdp, function(tau) 
  (1 + sum(W_sdp <= -tau)) / max(1, sum(W_sdp >= tau)))

alpha = 0.1
tau_hat_sdp <- taus_sdp[[which.min(FDP_hat_sdp[FDP_hat_sdp <= alpha]) |> names()]]
S_hat_sdp = which(W_sdp >= tau_hat_sdp)

print(paste("SDP tau:", tau_hat_sdp))
print(paste("Estimated FDP:", round((1 + sum(W_sdp <= -tau_hat_sdp)) / length(S_hat_sdp), 3)))

selected_indices_sdp <- which(W_sdp >= tau_hat_sdp)
selected_names_sdp <- colnames(X)[selected_indices_sdp]

print(paste("Number of variables selected by SDP:", length(selected_names_sdp)))
print(selected_names_sdp)
```

We obtain tau=2.4767, which returns an estimated FDP of 0.091 and selects 11 variables.
This plot better shows the results we obtained.

```{r}
cols_sdp <- ifelse(W_sdp >= tau_hat_sdp, "red", "black")

plot(Z_sdp[orig], Z_sdp[orig+p], col=cols_sdp, pch=19, asp=1, 
     xlab=expression(paste(lambda," when ",X[j]," enters ")),
     ylab=expression(paste(lambda," when ",tilde(X),"j enters ")),
     main=paste("SDP Selection (tau =", round(tau_hat_sdp, 3), ")"))
abline(a=0,b=1)
abline(h=tau_hat_sdp, lty=3, col="blue")
abline(v=tau_hat_sdp, lty=3, col="blue")
legend("topleft", legend=c("Selected", "Not Selected"), col=c("red", "black"), pch=19)
```


# Randomness inside knockoff

We will do 100 iterations of the knockoff procedure.
We decide to insert a variable into the final model if it gets selected more than 80% of the times.

```{r}
set.seed(42) 

M <- 100              
alpha <- 0.10        
cutoff <- 0.8       

selection_matrix <- matrix(0, nrow = ncol(X), ncol = M)
rownames(selection_matrix) <- colnames(X)
```

The cycle iterates the following steps: 

* Creates Xtilde (using SDP-Knockoff to maximise performance). By setting "Randomize" to TRUE, it finds a random plane generated by an orthonormal base of vectors which are orthogonal to the column space of X.
* Calculates the knockoff statistics.
* Finds tau such that FDP is lower than alpha.
* Stores the selected variables.

```{r}
for(i in 1:M) {
  
  Xtilde_random <- create.fixed(X, method = "sdp", randomize = TRUE)$Xk
  
  XX_loop <- cbind(X, Xtilde_random)
  fit_loop <- glmnet(XX_loop, y)
  indices_loop <- apply(fit_loop$beta, 1, first_nonzero)
  Z_loop = ifelse(is.na(indices_loop), 0, fit_loop$lambda[indices_loop] * n)
  W_loop = pmax(Z_loop[orig], Z_loop[orig+p]) * sign(Z_loop[orig] - Z_loop[orig+p])
  
  taus_loop = sort(c(0, abs(W_loop)))
  FDP_hat_loop = sapply(taus_loop, function(tau) 
    (1 + sum(W_loop <= -tau)) / max(1, sum(W_loop >= tau)))
  tau_hat_loop <- taus_loop[which(FDP_hat_loop <= alpha)[1]]
  
  if(!is.na(tau_hat_loop)) {
    selected_indices <- which(W_loop >= tau_hat_loop)
    
    if(length(selected_indices) > 0){
      selection_matrix[names(selected_indices), i] <- 1
    }
  }
  
  cat(paste("Iteration", i, ": Selected", sum(selection_matrix[,i]), "variables.\n"))
}
```

The stabilised variables are:

```{r}
probs <- rowMeans(selection_matrix)

stable.vars <- names(probs)[probs >= cutoff]

print(paste("Count:", length(stable.vars)))
print(stable.vars)
```

This graph represents, for each variable, the probability of being selected.

```{r}
plot(probs, type="h", lwd=2, 
     main="Stability Selection Frequency", 
     ylab="Probability of Selection", xlab="Variable Index")
abline(h=cutoff, col="red", lty=2)
```


# Random Forest

The Knockoff Filter is just a framework: it can be used with other variable selection techniques other than Lasso, such as Random Forests.
We will show an example, in which we'll use Variable Importance ("impurity") to build the Knockoff Statistic, and use SDP-Knockoffs to gain a better performance.

```{r}
set.seed(42)

XX_sdp_rf <- cbind(X, Xtilde_sdp) 

random_forest_importance <- function(X, y, ...) {
  df = data.frame(y=y, X=X)
  rfFit = ranger::ranger(y~., data=df, importance="impurity", write.forest=F, ...)
  as.vector(rfFit$variable.importance)
}


Z_rf = random_forest_importance(XX_sdp_rf, y) 


orig = 1:p
W_rf = abs(Z_rf[orig]) - abs(Z_rf[orig+p])
```

Initially, we select tau=0.001.

```{r}
tau_rf = 0.001

my_cols_rf <- ifelse(W_rf >= tau_rf, "red", "black")

plot(W_rf, col=my_cols_rf, type="h", lwd=2, xlab="Index j", 
     main="Random Forest Importance (SDP Knockoffs)", ylab="W statistic")
abline(h=tau_rf,lty=3)
abline(h=-tau_rf,lty=3)
```
For this selected tau, we obtain:

```{r}
S_hat_tau_rf = which(W_rf >= tau_rf)
(1 + sum(W_rf <= -tau_rf)) / length(S_hat_tau_rf)
```
An estimated FDP of 6.666, which means that for each significant variable, the model inserts more than 7 false positive discoveries.

Now, we select alpha=0.1 and the algorithm finds the best tau.

```{r}
taus = sort(c(0, abs(W_rf)))

FDP_hat_rf = sapply(taus, function(tau)
  (1 + sum(W_rf <= -tau)) / max(1, sum(W_rf >= tau)))

alpha = 0.1
tau_hat_rf <- taus[which(FDP_hat_rf <= alpha)[1]]

S_hat_rf = which(W_rf >= tau_hat_rf)
print(paste("Selected by RF:", length(S_hat_rf)))
print(colnames(X)[S_hat_rf])

(1 + sum(W_rf <= -tau_hat_rf)) / length(S_hat_rf)
```
We see how the algorithm doesn't find any combination of variables that could grant an estimate of FDP lower than 0.1.
This result can be express through this graph, that shows the FDP estimate for each possible tau:

```{r}
plot(taus, FDP_hat_rf, type="l", lwd=2, xlab=expression(tau), ylab="FDP Estimate",
     main="RF Threshold Selection")
abline(h=alpha, col="red", lty=2)
abline(v=tau_hat_rf, col="blue", lwd=2)
```

We can see how this model doesn't work with Fixed-X knockoff. However, this can be explained; Random Forest is not a linear model, because it places great importance on the true distribution of data. Since our predictors were fixed, its assumptions were violated and results lacked.
This is one of the main reason Model-X Knockoffs are used.

# Model Evaluation

We will compare the performances of 4 different models:
* a simple linear regression
* a LASSO regression
* a linear regression with equi-knockoff selected variables
* a linear regression with randomized SDP-knockoff selected variables

Firstly, the models are fit on the training data:

```{r}
rm(list = setdiff(ls(all = F), c("df.train", "df.test", "equi.vars", "stable.vars")))

X <- df.train |> select(-SalePrice) |> as.matrix()
y <- df.train$SalePrice

mod.lm <- lm(SalePrice ~ ., data = df.train)
mod.lasso <- glmnet(X, y); l <- cv.glmnet(X, y)$lambda.min
mod.ko.equi <- lm(SalePrice ~ ., data = df.train[, c(equi.vars, "SalePrice")])
mod.ko.rand <- lm(SalePrice ~ ., data = df.train[, c(stable.vars, "SalePrice")])
```

Then RMSE is evaluated on the test data:

```{r}
X <- df.test |> select(-SalePrice) |> as.matrix()
y <- df.test$SalePrice

y.lm <- predict(mod.lm, newdata = df.test) |> exp()
y.lasso <- predict(mod.lasso, newx = X, s = l) |> exp()
y.ko.equi <- predict(mod.ko.equi, newdata = df.test[, c(equi.vars, "SalePrice")]) |> exp()
y.ko.rand <- predict(mod.ko.rand, newdata = df.test[, c(stable.vars, "SalePrice")]) |> exp()
y <- exp(y)

c(linear = RMSE(y.lm, y), 
  lasso = RMSE(y.lasso, y), 
  equicorr = RMSE(y.ko.equi, y), 
  randsdp = RMSE(y.ko.rand, y))
```
The results highlight the superiority of the SDP-knockoff selected variables, which have a considerable edge over the other methods. Interestingly, the equicorrelated knockoffs do not show an increase in performance over a basic LASSO regression.
